Exploring the tree of numerical semigroups

### 2014-10-09 Papier/article.tex
Florent's rereading

Here we describe a new algorithm for the exploration of the tree of numerical
semigroup $\mathcal{T}$ and achieve the computation of $n_g$ for $g\leq 67$.
The cornerstone of our method is a new combinatorial representation of
numerical semigroup that is well suited for exploration of the
tree~$\mathcal{T}$ and allow large code optimization essentially based on use of
vectorial instructions and parallelization.
The goal of the paper is twofold:
first to present our new encoding of the semigroup and the associated
algorithms, and second to present the optimization techniques which allows for
those kinds of algorithms to get speedups by factors of hundreds and even
thousands. We claim that these technique are fairly general for those kinds of
algorithms. As a support for the claim, we applied it to an algorithm of
N.~Borie enumerating integer vector modulo permutation groups~\cite{Borie} and
got a speedup by a factor larger than 2000 using 8 cores !

@@

Finally we test our algorithm on a machine holding two 
\tt{Intel\texttrademark{} Xeon\texttrademark{} X5650} CPU running 
at \texttt{2.67GHz}.
Each of those CPU have 6 physical cores that are able to run 12 threads : the 
machine has also 12 physical cores.  
For more than $12$ threads, the speedup should be much less when
adding more core. The following table resume the time needed by the algorithm 
to explore $\mathcal{T}_{50}$ on diffenent number of threads (the double 
vertical line delimites the use of \texttt{Hyper Threadind}):
\begin{figure}[h!]
\begin{tabular}{|c|c|c|c|c|c||c|c|c|}
\hline
 Threads & 1 & 2 & 4 & 8 & 12 & 16 & 20 & 24\\
 \hline
 Time (s) & 3709 & 1865 & 932.4 & 486.8 & 325.7 & 311.2 & 302.3 & 290.2\\
 \hline
\end{tabular}
\end{figure}
% 
% We can see that this is indeed happening under the
% 
% horizontal bar in the following table.
%   \begin{figure}[h!]
% \begin{tabular}{|c|c|c|c|c|c|}
% \hline
% Threads & 30 & 35 & 40 & 45 & 50 \\
% \hline
% 1 &  0.33 & 2.50 & 26.1 & 307 & 3709 \\
% 2 &  0.17 & 1.25 & 13.05 & 153.5 & 1865\\
% 3 &  0.11 & 0.71 & 8.57 & 102.6 & 1243 \\
% 4 &  0.10 & 0.55 & 6.45 &  77.21 & 932.4 \\
% 8 &  0.10 & 0.36 & 3.55 &  40.22 & 486.8 \\
% 12 &  0.10 & 0.36 & 2.61 & 27.34 & 325.7 \\\hline
% 16 &  0.10 & 0.37 & 2.50 & 25.96 & 311.2 \\
% 20 &  0.10 & 0.37 & 2.69 & 26.04 & 302.3 \\
% 24 &  0.12 & 0.51 & 3.18 & 25.27 & 290.2 \\
% \hline
% \end{tabular}
% \end{figure}

@@

XXX

>>>>

Here we describe a new algorithm for the exploration of the tree of numerical
semigroup $\mathcal{T}$ and achieve the computation of $n_g$ for $g\leq 67$.
The cornerstone of our method is a new combinatorial representation of
numerical semigroup that is well suited for exploration of the
tree~$\mathcal{T}$ and allow large code optimization essentially based on use of
vectorial instructions and parallelization.
The goal of the paper is twofold:
first to present our new encoding of the semigroup and the associated
algorithms, and second to present the optimization techniques which allows, for
those kinds of algorithms, to get speedups by factors of hundreds and even
thousands. We claim that these technique are fairly general for those kinds of
algorithms. As a support for the claim, we applied it to an algorithm of
N.~Borie enumerating integer vector modulo permutation groups~\cite{Borie} and
got a speedup by a factor larger than 2000 using 8 cores !

@@

Finally we test our algorithm on a machine holding two 
\tt{Intel\texttrademark{} Xeon\texttrademark{} X5650} CPU running 
at \texttt{2.67GHz}.
Each of those CPU have 6 physical cores that are able to run 12 threads : the 
machine has also 12 physical cores.  
For more than $12$ threads, the speedup should be much less when
adding more core. The following table resumes the time needed by the algorithm 
to explore $\mathcal{T}_{50}$ on different number of threads (the double 
vertical line delimites the use of \texttt{Hyper Threadind}). For reference,
we put in the column called ``C++'' the computation time of the same
program compiled without Cilk.
\begin{figure}[h!]
\begin{tabular}{|c|c|c|c|c|c|c||c|c|c|}
\hline
 Threads & C++ & 1 & 2 & 4 & 8 & 12 & 16 & 20 & 24\\
 \hline
 Time (s) &      & 3709 & 1865 & 932.4 & 486.8 & 325.7 & 311.2 & 302.3 & 290.2\\
 \hline
\end{tabular}
\end{figure}

% 
% We can see that this is indeed happening under the
% 
% horizontal bar in the following table.
%   \begin{figure}[h!]
% \begin{tabular}{|c|c|c|c|c|c|}
% \hline
% Threads & 30 & 35 & 40 & 45 & 50 \\
% \hline
% 1 &  0.33 & 2.50 & 26.1 & 307 & 3709 \\
% 2 &  0.17 & 1.25 & 13.05 & 153.5 & 1865\\
% 3 &  0.11 & 0.71 & 8.57 & 102.6 & 1243 \\
% 4 &  0.10 & 0.55 & 6.45 &  77.21 & 932.4 \\
% 8 &  0.10 & 0.36 & 3.55 &  40.22 & 486.8 \\
% 12 &  0.10 & 0.36 & 2.61 & 27.34 & 325.7 \\\hline
% 16 &  0.10 & 0.37 & 2.50 & 25.96 & 311.2 \\
% 20 &  0.10 & 0.37 & 2.69 & 26.04 & 302.3 \\
% 24 &  0.12 & 0.51 & 3.18 & 25.27 & 290.2 \\
% \hline
% \end{tabular}
% \end{figure}

@@

Bureau 33, Laboratoire de Recherche en Informatique (UMR CNRS 8623)

Bâtiment 650, Universit\'e Paris Sud 11, 91405 Orsay CEDEX, France
